# We start by using elements of the Inference vignette

```{r}
library(pander)
library(dplyr)
library(tidyr)
library(fluEvidenceSynthesis)
library(reshape2)
library(ggplot2)
library(reshape2)
```

## Load all of the relevant data for the inference process

### Two age groups, single risk group

```{r}
# Here we load the needed data and regroup it from 5 into two age groups.
data("demography")
data("polymod_uk")
data("ili")

# Virologically confirmed infections
# THIS IS A KEY PIECE WHICH WE MAY BE MISSING
data("confirmed.samples")

# Group all the example data into two age groups:
polymod <- polymod_uk[,c(1,2)]
polymod[,3] <- rowSums(polymod_uk[,c(3,4,5,6,7,8)])
polymod[,4] <- polymod_uk[,9]

ili_df <- ili
ili_df$ili <- ili$ili[,c(1,2)]
ili_df$ili[,1] <- rowSums(ili$ili[,c(1,2,3,4)])
ili_df$ili[,2] <- ili$ili[,5]
ili_df$total.monitored <- ili$total.monitored[,c(1,2)]
ili_df$total.monitored[,1] <- rowSums(ili$total.monitored[,c(1,2,3,4)])
ili_df$total.monitored[,2] <- ili$total.monitored[,5]
```

The below is the [**problematic bit**]{.underline}, loading confirmed sample numbers (virological/serological data).

```{r}
confirmed.samples_df <- confirmed.samples
confirmed.samples_df$positive <- confirmed.samples$positive[,c(1,2)]
confirmed.samples_df$positive[,1] <- rowSums(confirmed.samples$positive[,c(1,2,3,4)])
confirmed.samples_df$positive[,2] <- confirmed.samples$positive[,5]
confirmed.samples_df$total.samples <- confirmed.samples$total.samples[,c(1,2)]
confirmed.samples_df$total.samples[,1] <- rowSums(confirmed.samples$total.samples[,c(1,2,3,4)])
confirmed.samples_df$total.samples[,2] <- confirmed.samples$total.samples[,5]
```

#### Load coverage data (historical)

```{r}
data(coverage)
# Coverage rates for the low risk <65 and low risk 65+
cov <- coverage[,c("Under.65","X65")]/100.0
```

#### Create vaccination calendar object

```{r}
vaccine_calendar <- as_vaccination_calendar(efficacy = c(0.7, 0.4), 
                                        dates = coverage$Date,
                                        coverage = cov, 
                                        no_age_groups = 2, no_risk_groups = 1)
```

#### Choose initial parameter guesses (appropriate for two age groups, single risk group, hence 7 entries)

```{r}
initial_parameters <- c(0.1, 0.1, 1e-5, 0.16, 0.5, 0.5, -0.15) # 7-vector
names(initial_parameters) <- c("epsilon_1", "epsilon_2", "psi", "transmissibility", "susceptibility_1", "susceptibility_2", "initial_infected")
```

#### Run inference (two age groups, single risk group)

Best to use the in-built inference function, quite honestly.

```{r}
inference.res <- inference(demography = demography,
                      vaccine_calendar = vaccine_calendar,
                      polymod_data = as.matrix(polymod),
                      ili = ili_df$ili,
                      mon_pop = ili_df$total.monitored,
                      n_pos = confirmed.samples_df$positive,
                      n_samples = confirmed.samples_df$total.samples,
                      initial = initial_parameters,
                      age_groups = c(65), # NOTICE THIS, SEE DOCS FOR DETAILS
                      nbatch = 3000,
                      nburn = 1000, blen = 50)
```

Now we define the **custom inference** function. Note that this function is an absolute pain. It has a lot of things that are simply hard-coded and make it very very difficult to use/**edit**.

```{r}
# The custom inference function. In this example the custom inference function 
# performs exactly the same inference as the original C++ function (above). 
# It is up to the user to change this in a way that works for their analysis.
custom_inference <- function(demography, vaccine_calendar, polymod_data, ili, 
                             mon_pop, n_pos, n_samples, initial, mapping,
                             nbatch, nburn, blen) {
  current.contact.ids <- seq(1,nrow(polymod_uk))
  proposed.contact.ids <- current.contact.ids
  
  # Seven age groups used in the model
  age.group.limits <- c(1,5,15,25,45,65)
  
  # Sum all populations with a certain age into their corresponding age group
  age.group.sizes.5 <- stratify_by_age(demography, c(5,15,45,65))
  
  if (missing(mapping))
    mapping <- age_group_mapping(age.group.limits, c(5,15,45,65))
  
  # Define the actual log likelihood function
  llikelihood <- function( pars ) {
    # Resample contact ids 
    proposed.contact.ids <<- current.contact.ids
    if (runif(1,0,1) < 0.1) {
      rs <- round(runif(2,1,length(proposed.contact.ids)))
      proposed.contact.ids[rs[1]] <<- rs[2]
    }
    
    contacts <- contact_matrix(as.matrix(polymod_uk[proposed.contact.ids,]),
                               demography, age.group.limits )
    
    age.groups <- stratify_by_age(demography, 
                                  age.group.limits )
    
    # Fraction of each age group classified as high risk
    # We can classify a third risk group, but we are not doing
    # that here (the second row is 0 in our risk.ratios matrix)
    risk.ratios <- matrix(c(
      0.021, 0.055, 0.098, 0.087, 0.092, 0.183, 0.45, 
      0, 0, 0, 0, 0, 0, 0                          
    ), ncol = 7, byrow = T)
    
    # Population sizes in each age and risk group
    popv <- stratify_by_risk(
      age.groups, risk.ratios);
    
    # Population size initially infected by age and risk group
    initial.infected <- rep( 10^pars[9], 7 ) # 7 for number of age groups (in the ILI data, not virological)
    initial.infected <- stratify_by_risk(
      initial.infected, risk.ratios );
    
    # Run simulation
    # Note that to reduce complexity 
    # we are using the same susceptibility parameter for multiple age groups
    odes <- infectionODEs( popv, initial.infected,
                           vaccine_calendar,
                           contacts,
                           c(pars[6], pars[6], pars[6],
                             pars[7], pars[7], pars[7], pars[8]),
                           transmissibility = pars[5],
                           c(0.8,1.8), 7 )
    
    # Ignore times row
    odes <- odes[,2:22]
    
    # Convert the 7 age groups for each risk group to 5 groups
    from <- as.numeric(mapping$from)
    to <- as.numeric(mapping$to)
    converted.odes <- matrix(0, nrow = nrow(odes), ncol = max(to))
    for (i in 1:nrow(mapping)) {
      # all three age groups
      fv <- c(0,7,14) + from[i]
      converted.odes[,to[i]] <- converted.odes[,to[i]] + mapping$weight[i]*rowSums(odes[,fv]) 
    }
    
    # For each week and each group sum log likelihood
    epsilons <- c(pars[1], pars[1], pars[2], pars[2], pars[3])
    ll <- log_likelihood_cases(
      epsilons,pars[4], as.matrix(converted.odes),
      age.group.sizes.5, ili, mon_pop,
      n_pos, n_samples)
    return(ll)
  }
  llprior <- function(pars) {
    if (any(pars[1:8] < 0) || any(pars[1:4] > 1) || any(pars[6:8] > 1)
        || pars[9] < log(0.00001) || pars[9] > log(10) )
      return(-Inf)
    
    lprob <- dnorm(pars[5], 0.1653183, 0.02773053, 1)
    lprob <- lprob + dlnorm(pars[1], -4.493789, 0.2860455, 1)
    lprob <- lprob + dlnorm(pars[2], -4.117028, 0.4751615, 1)
    lprob <- lprob + dlnorm(pars[3], -2.977965, 1.331832, 1)
    
    return(lprob)
  }
  
  # Store the contact ids used during inference
  contact.ids <- list()
  
  # Run adaptive.mcmc
  mcmc.result <- adaptive.mcmc(lprior = llprior, llikelihood = llikelihood, 
                               outfun = function() { 
                                 contact.ids[[length(contact.ids)+1]] <<-  current.contact.ids
                               },
                               acceptfun = function() {
                                 current.contact.ids <<- proposed.contact.ids
                               },
                               nburn = nburn, 
                               initial = initial,
                               nbatch = nbatch, blen = blen)
  mcmc.result$contact.ids <- t(data.frame(contact.ids))
  mcmc.result
}
```

The resulting custom_inference function can be called similarly to the original inference function, but it's probably best not to...

```{r}
# The resulting custom inference function can be called similarly to the original inference function.

# initial.parameters <- c(0.01188150, 0.01831852, 0.05434378, 1.049317e-05, 0.1657944, 0.3855279, 0.9269811, 0.5710709, -0.1543508)
# 
# # Adding names for clarity, is not actually needed
# names(initial.parameters) <- c("espilon_1", "epsilon_2", "epsilon_3", "psi",                               "transmissibility", "susceptibility_1", "susceptibility_2", "suceptibility_3","initial_infected")
# 
# 
# inference.res <- custom_inference(demography = demography,
#                       vaccine_calendar = vaccine_calendar,
#                       polymod_data = as.matrix(polymod_uk),
#                       ili = ili$ili,
#                       mon_pop = ili$total.monitored,
#                       n_pos = confirmed.samples$positive,
#                       n_samples = confirmed.samples$total.samples,
#                       initial = initial.parameters,
#                       nbatch = 1000,
#                       nburn = 1000, blen = 5 )

```

# Analysing the results

Plotting the resulting posterior parameter values[\^vignette-blend-1].

We need to change the names used here in colnames depending on what number of parameters we have inferred!

```{r,cache=T,fig.width=5,fig.height=5}

# Just below is the case for a 9-vector of parameters, based on the more complicated example in the online inference vignette
# colnames(inference.res$batch) <- c("eps1", "eps2", "eps3", "psi", "q",
#                                        "susc1", "susc2", "susc3", "I0")

# Just below is the case for a 7-vector of parameters, based on the more complicated example in the online inference vignette
colnames(inference.res$batch) <- c("eps1", "eps2", "psi", "q",
                                        "susc1", "susc2", "I0")

ggplot(data=melt(inference.res$batch)) + facet_wrap( ~ Var2, ncol=3, scales="free" ) + geom_histogram(aes(x=value), bins=25)

```
