# We start by using elements of the Inference vignette

```{r}
library(pander)
library(dplyr)
library(tidyr)
library(fluEvidenceSynthesis)
library(reshape2)
library(ggplot2)
library(reshape2)
```

## Load all of the relevant data for the inference process

### Two age groups, single risk group

```{r}
# Here we load the needed data and regroup it from 5 into two age groups.
data("demography")
data("polymod_uk")
data("ili")

# Virologically confirmed infections
# THIS IS A KEY PIECE WHICH WE MAY BE MISSING
data("confirmed.samples")

# Group all the example data into two age groups:
polymod <- polymod_uk[,c(1,2)]
polymod[,3] <- rowSums(polymod_uk[,c(3,4,5,6,7,8)])
polymod[,4] <- polymod_uk[,9]

ili_df <- ili
ili_df$ili <- ili$ili[,c(1,2)]
ili_df$ili[,1] <- rowSums(ili$ili[,c(1,2,3,4)])
ili_df$ili[,2] <- ili$ili[,5]
ili_df$total.monitored <- ili$total.monitored[,c(1,2)]
ili_df$total.monitored[,1] <- rowSums(ili$total.monitored[,c(1,2,3,4)])
ili_df$total.monitored[,2] <- ili$total.monitored[,5]
```

The below is the [**problematic bit**]{.underline}, loading confirmed sample numbers (virological/serological data).

```{r}
confirmed.samples_df <- confirmed.samples
# Create matrices of zeros with the same dimensions as the original matrices
# For confirmed.samples$positive, assuming it has two columns after simplification
confirmed.samples_df$positive <- matrix(0, nrow = nrow(confirmed.samples$positive), ncol = 2)

# For confirmed.samples$total.samples, also assuming it has two columns after simplification
confirmed.samples_df$total.samples <- matrix(0, nrow = nrow(confirmed.samples$total.samples), ncol = 2)
```

#### Load coverage data (historical)

```{r}
data(coverage)
# Coverage rates for the low risk <65 and low risk 65+
cov <- coverage[,c("Under.65","X65")]/100.0
```

#### Create vaccination calendar object

```{r}
vaccine_calendar <- as_vaccination_calendar(efficacy = c(0.7, 0.4), 
                                        dates = coverage$Date,
                                        coverage = cov, 
                                        no_age_groups = 2, no_risk_groups = 1)
```

#### Choose initial parameter guesses (appropriate for two age groups, single risk group, hence 7 entries)

```{r}
initial_parameters <- c(0.1, 0.1, 1e-5, 0.16, 0.5, 0.5, -0.15) # 7-vector
names(initial_parameters) <- c("epsilon_1", "epsilon_2", "psi", "transmissibility", "susceptibility_1", "susceptibility_2", "initial_infected")
```

#### Run inference (two age groups, single risk group)

Best to use the in-built inference function, quite honestly.

```{r}
inference.res <- inference(demography = demography,
                      vaccine_calendar = vaccine_calendar,
                      polymod_data = as.matrix(polymod),
                      ili = ili_df$ili,
                      mon_pop = ili_df$total.monitored,
                      n_pos = confirmed.samples_df$positive,
                      n_samples = confirmed.samples_df$total.samples,
                      initial = initial_parameters,
                      age_groups = c(65), # NOTICE THIS, SEE DOCS FOR DETAILS
                      nbatch = 3000,
                      nburn = 1000, blen = 50)
```

# Analysing the results

Plotting the resulting posterior parameter values[\^vignette-blend-1].

We need to change the names used here in colnames depending on what number of parameters we have inferred!

```{r}

# Just below is the case for a 9-vector of parameters, based on the more complicated example in the online inference vignette
# colnames(inference.res$batch) <- c("eps1", "eps2", "eps3", "psi", "q",
#                                        "susc1", "susc2", "susc3", "I0")

# Just below is the case for a 7-vector of parameters, based on the more complicated example in the online inference vignette
colnames(inference.res$batch) <- c("eps1", "eps2", "psi", "q",
                                        "susc1", "susc2", "I0")

ggplot(data=melt(inference.res$batch)) + facet_wrap( ~ Var2, ncol=3, scales="free" ) + geom_histogram(aes(x=value), bins=25)

```



```{r}
library(ggplot2)
library(reshape2)

# Assuming inference.res$batch is already defined and colnames set
melted_data <- melt(inference.res$batch)

# Improved plot with adjusted labels and theme
ggplot(melted_data, aes(x=value)) + 
  geom_histogram(bins=25) + 
  facet_wrap(~ Var2, ncol=3, scales="free") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8), # Rotate x-axis labels and adjust size
        axis.text.y = element_text(size = 8), # Adjust y-axis label size
        strip.text = element_text(size = 8)) + # Adjust facet label size
  labs(x = NULL, y = "Frequency") + # Optionally remove x label for cleanliness
  theme_minimal() # Use a minimal theme for more space



```

